---
fontsize: 11pt
geometry: margin=0.20in
output:
  pdf_document:
    fig_height: 3
    fig_width: 4
    latex_engine: xelatex
---
### Analysis of ToothGrowth data

Here is a basic summary of the data (ToothGrowth dataset) as well as the data aggregated by supplement and dose, showing there are 60 animals in a factorial design by supplement and dose.  In other words, 30 animals assigned to receive orange juice and 30 to vitamin C, each recieving one of three doses of the supplement (a 10 animal x 6 factor design):
```{r, echo=FALSE}
library(datasets)
data(ToothGrowth)
ToothGrowth$dose=as.factor(ToothGrowth$dose)
```
```{r, echo=FALSE}
summary(ToothGrowth)
```
```{r, echo=FALSE}
aggMean=aggregate(len~supp+dose, data=ToothGrowth, mean)
aggSD=aggregate(len~supp+dose, data=ToothGrowth, sd)
aggMean
```

I will use a t-test to make a 95% confidence interval for the mean difference in tooth growth depending on supplement type. The assumptions here are: that this is an unpaired mean differences test so there is no other cause of tooth length differences other than supplement (the guninea pigs are assumed to have same environment, diet, genetic background etc...);the populations are approximately normal; that the variances (of the populations) are approximately equal; the samples are independent:
```{r, echo=FALSE}
suppOJ=subset(ToothGrowth, supp=="OJ")
suppVC=subset(ToothGrowth, supp=="VC")
t.test(suppOJ$len,suppVC$len,paired=F, var.equal=T)
```
Assuming the null hypothesis is that there is no difference (difference=zero), we would not reject the null hypothesis in this case as the confidence interval includes zero, p=0.06, 95% CI: (-0.17, 7.57). Orange juice appears to be a better supplement for tooth growth than vitamin C with a mean difference of 3.7, but this does not quite reach statistical significance.

For the dose experiment, I will do three t-tests comparing the three doses against eachother, two at a time.  I will use the same assumptions here:
```{r, echo=FALSE}
dose0.5 = subset(ToothGrowth,dose==0.5)
dose1.0 = subset(ToothGrowth,dose==1.0)
dose2.0 = subset(ToothGrowth,dose==2.0)
t.test(dose1.0$len, dose0.5$len, paired=F, var.equal=T)
t.test(dose2.0$len, dose0.5$len, paired=F, var.equal=T)
t.test(dose2.0$len, dose1.0$len, paired=F, var.equal=T)
```
Here, the p-value for each test is very small, indicating a definite increase in growth with increasing dose.  The 95% confidence interval for this ranges from (12.83, 18.16) comparing dose 2.0 to 0.5 dose, to (3.73, 8.99) comparing dose  2.0 to 1.0, though they are all above the zero that would be predicted under the null hypothesis.

However, the above values assume there is no interaction between supplement and dose, but a plot of the aggreagted data shows this is not true. This is one major limitation of considering the supplements and dose as independent variables. It appears that OJ and Vitamin C work equally well at the maximum dose, though OJ is superior at the lower doses, as shown in a side-by-side plot of the aggregated data:

```{r, echo=FALSE}
library(ggplot2)
ggplot(data=aggMean,aes(x=as.factor(dose), y=len,fill=supp))+geom_bar(stat="identity",position=position_dodge())+theme_bw()+labs(title="Toothgrowth vs supplement and dose", x="Dose",y="Tooth Length")
```

The full code for this project is available at: https://github.com/Allison-Miller/Statistical-Inference/blob/master/ToothGrowth.Rmd
